RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-01-21 gmas.tex
update

\subsubsection{Irrelevant actor}

The transcript excerpt of the first example is shown in table~\ref{table:transcript:irrelevant-actor} in the appendix. It consists of participant P1 putting forth the suggestion to include actor \emph{Development team} into the model. This is then questioned by person P2, who argues that the professor will develop the software, so there won't be any development team.

We formalized the first statement as an instantiation of argument scheme AS0: \emph{Actor development team is relevant}. This argument is then attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}. This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0. This is shown in figure~\ref{fig:examples:relevant-actor:arguments}.

Argument scheme AS0 is an argument to include actor \emph{Development team} into the GRL model. Therefore, by attacking it the argument is no longer accepted, and as a result the actor is disabled in the resulting GRL model (figure~\ref{fig:examples:relevant-actor:grl}). Note we added a red traceability link from the rejected argument to the disabled GRL element.

\begin{figure}[ht!]
\centering
\begin{subfigure}[b]{0.22\textwidth}
        \centering
        \begin{tikzpicture}[->]
        \node (a0) [shape=rectangle, draw=black!60, text=black!40] at (0,0) {
            \begin{tabularx}{\textwidth}{X}
                \textbf{(AS0)} Development team is relevant 
            \end{tabularx}
        } ;
        \node (a1) [shape=rectangle,draw] at (0,-2){
            \begin{tabularx}{\textwidth}{X}
                \textbf{(CQ0)} Development team is not relevant (\emph{The professor makes the software})
            \end{tabularx}
        };
        \node[circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt] (trace) at (1.2,-0.3) {};

         \path[line width=1mm]
    (a1) edge[draw=black] (a0);
\end{tikzpicture}
        \caption{Arguments}
        \label{fig:examples:relevant-actor:arguments}
    \end{subfigure}\hfill
    \begin{subfigure}[b]{0.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{img/actor_disabled}
        \caption{GRL elements}
        \label{fig:examples:relevant-actor:grl}
    \end{subfigure}
\caption{Argument schemes and critical questions (left) and GRL model (right) of a discussion about the relevance of actor Development team.}
\label{fig:examples:relevant-actor}
\end{figure}

>>>>

\subsubsection{Example 1: Introducing a new element (INTRO)}

\subsubsection{Example 2: Disabling an element (DISABLE)}

\subsubsection{Example 3: Change decomposition type (REPLACE)}

\subsubsection{Example 4: Clarifying an element (REPLACE)}

The transcript excerpt of the second example is shown in table~\ref{table:transcript:task-clarification} in the appendix and comes from transcript $t_3$. It consists of a number of clarification steps, resulting in the task \emph{Choose a road pattern}.

\begin{figure}[ht!]
\centering
        \begin{tikzpicture}[->]
        \node (a0) [shape=rectangle, draw=black!60, text=black!40] at (-2,0) {
            \begin{tabularx}{0.2\textwidth}{X}
                \textbf{(AS2)} Actor \emph{student} has task \emph{create road}
            \end{tabularx}
        };
        \node (a1) [shape=rectangle, draw=black!60, fill=white, text=black!40] at (-2,-2) {
            \begin{tabularx}{0.2\textwidth}{X}
                \textbf{(AS2)} Actor \emph{student} has task \emph{choose a pattern}
            \end{tabularx}
        };
        \node (a2) [shape=rectangle, draw=black!60, fill=white, text=black!40] at (-2,-4.2) {
            \begin{tabularx}{0.2\textwidth}{X}
                \textbf{(AS2)} Actor \emph{student} has task \emph{choose a pattern preference}
            \end{tabularx}
        };
        \node (a3) [shape=rectangle] at (-2,-6.7) {
            \begin{tabularx}{0.2\textwidth}{X}
                \textbf{(AS2)} Actor \emph{student} has task \emph{choose a road pattern}
            \end{tabularx}
        };
        \node[draw=none] (actor) at (2.3,-4) { \includegraphics[scale=0.5]{img/task_choose_road_pattern} };
        \node[circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt] (trace1) at (-.5,-6.5) {};
        \node[circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt] (trace2) at (2.2,-4.4) {};

\begin{pgfonlayer}{background}
         \path[line width=1mm]
    (a1) edge[draw=black] (a0)
    (a2) edge[draw=black] (a1)
    (a2) edge[draw=black, bend right=20] (a0)
    (a3) edge[draw=black] (a2)
    (a3) edge[draw=black, bend right=20] (a1)
    (a3) edge[draw=black, bend right=40] (a0);
\end{pgfonlayer}
     \path[line width=0.5mm, dotted]
    (a0) edge [bend right=50] node [left,draw=none] {CQ12a} (a1)
    (a1) edge [bend right=50] node [left,draw=none] {CQ12a} (a2)
    (a2) edge [bend right=50] node [left,draw=none] {CQ12a} (a3);

    \path[draw, densely dotted,-,line width=0.5mm] (trace1) edge[draw=green] (trace2);
\end{tikzpicture}
\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the clarification of a task.}
\label{fig:examples:clarification}
\end{figure}

The formalized argument schemes and critical questions are shown in figure~\ref{fig:examples:clarification}. The discussion starts with the first instantiation of argument scheme AS2: \emph{Actor student has task create road}. This argument is then challenged with critical question CQ12a: \emph{Is the task create road clear?}. Answering this question results in a new instantiation of argument scheme AS2: \emph{Actor student has task choose a pattern}. This process is repeated three times, resulting in the final argument scheme \emph{Actor student has task choose a road pattern}. This final argument is unattacked and has a corresponding intentional element (right image).

What is clearly shown in this example is that a clarification step does not only attacks the previous argument, but it also attacks all arguments previously used to describe the element. For instance, the final argument on the bottom of figure~\ref{fig:examples:clarification} attacks all previous arguments. If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked.

\subsubsection{Example 5: Generic counterargument (ATTACK)}

\todo{Marc}{Marc}{Add the additional counterargument and add to critical questions}

The transcript excerpt of the first example is shown in table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$. It consists of participant P1 putting forth the suggestion to include actor \emph{Development team} into the model. This is then questioned by person P2, who argues that the professor will develop the software, so there won't be any development team.

\begin{figure}[ht!]
\centering
        \begin{tikzpicture}[->]
        \node (a0) [shape=rectangle, draw=black!60, text=black!40] at (-2,0) {
            \begin{tabularx}{0.2\textwidth}{X}
                \textbf{(AS0)} Development team is relevant 
            \end{tabularx}
        } ;
        \node (a1) [shape=rectangle,draw] at (-2,-2.5){
            \begin{tabularx}{0.2\textwidth}{X}
                Development team is not relevant (\emph{The professor makes the software})
            \end{tabularx}
        };
        \node[draw=none] (actor) at (2.3,-1) { \includegraphics[scale=0.8]{img/actor_disabled} };
        \node[circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt] (trace1) at (-.5,-.3) {};
        \node[circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt] (trace2) at (2.2,-0.1) {};

         \path[line width=1mm]
    (a1) edge[draw=black] (a0);
    \path[line width=0.5mm, dotted]
    (a0) edge [bend right=50] node [left,draw=none] {CQ0} (a1);
    \path[draw, densely dotted,-,line width=0.5mm] (trace1) edge[draw=red] (trace2);
\end{tikzpicture}
\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of a discussion about the relevance of actor Development team.}
\label{fig:examples:relevant-actor}
\end{figure}

We formalized the first statement as an instantiation of argument scheme AS0: \emph{Actor development team is relevant}. This argument is then attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}. This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0. This is shown in figure~\ref{fig:examples:relevant-actor}, left image.

Argument scheme AS0 is an argument to include actor \emph{Development team} into the GRL model. Therefore, by attacking it the argument is no longer accepted, and as a result the actor is disabled in the resulting GRL model (figure~\ref{fig:examples:relevant-actor}, right image). Note we added a red traceability link from the rejected argument to the disabled GRL element.

### 2017-01-21 transcripts.tex
update

\begin{table}[!htbp]
\centering
\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
\hline
0:11:00.8 (P2) & Development team, I don't know. Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
\hline
\end{tabular}
\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
\label{table:transcript:as0-cq0}

\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
\hline
0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
\hline
\end{tabular}
\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
\label{table:transcript:as1-as8}

>>>>

\begin{table}[!htbp]
\centering
\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
\hline
0:11:00.8 (P2) & Development team, I don't know. Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
\hline
\end{tabular}
\caption{Discussion about the relevance of an actor (transcript $t_3$)}
\label{table:transcript:irrelevant-actor}

\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:17:39.5 (P1) & And in that process there are activities like create a visual map, create a road& \textbf{[14 task (AS2)]} Student has task ``Create road''\\
\hline
0:24:36.0 (P3)	& And, well interaction. Visualization sorry. Or interaction, I don't know. So create a visual map would have laying out roads and a pattern of their choosing. So this would be first, would be choose a pattern. & \multirow{3}{60mm}{\textbf{[31 critical question CQ?? for 14]} Is Task ``Create road'' clear?\\
\textbf{[32 answer to 31]} no, according to the specification the student should choose a pattern.\\
\textbf{[32a REPLACE]} ``Create road'' becomes ``Choose a pattern''}\\
\cline{1-2}
0:24:55.4 (P1) &	How do you mean, choose a pattern	&\\
\cline{1-2}
0:24:57.5 (P3)	& Students must be able to create a visual map of an area, laying out roads in a pattern of their choosing	&\\
\hline
0:25:07.5 (P1)	& Yeah I'm not sure if they mean that. I don't know what they mean by pattern in this case. I thought you could just pick roads, varying sizes and like, broads of roads. & \textbf{[33 critical question CQ?? for 32a]} Is ``Choose a pattern'' clear?\newline
\textbf{[34 answer to 33]} No, not sure what they mean by a pattern.\\
\hline
0:25:26.0 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \multirow{3}{60mm}{\textbf{[34a REPLACE]} ``Choose a pattern'' becomes ``Choose a pattern preference''}\\
\cline{1-2}
0:25:36.3 (P1) & Yeah, selecting a- I don't know &\\
\cline{1-2}
0:25:38.0 (P3)	& Pattern preference maybe? As in, maybe we can explain this in the documentation &\\
\hline
0:25:43.9 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ?? for 34a]} Is ``Choose a pattern preference'' clear?\newline
\textbf{[36 answer to 35]} no, what kind of pattern?\\
\hline
0:25:47.4 (P3) & Maybe, I don't know it's- & \multirow{2}{60mm}{\textbf{[36a rename]} ``Choose a pattern preference'' becomes ``Choose a road pattern''}\\
\cline{1-2}
0:25:48.5 (P1)	& [inaudible] a road pattern&\\
\hline
\end{tabular}
\caption{Clarifying the name of a task (transcript $t_3$)}
\label{table:transcript:task-clarification}
\end{table}

\begin{table}[!htbp]
\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
\hline
0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
\hline
\end{tabular}
\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
\label{table:transcript:as1-as8}
